---
title: "Functionalizing Rasters from NEON"
author: "R. Remigio"
date: "June 21, 2016"
output: pdf_document
---

## Objectives

Start by putting up objectives/tasks that students will be working though:
1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

1. For the CHM, set values == 0 to NA (not trees)

1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that 6m is generally grasses / understory, 10m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize density and plot vertical cutoff lines.

1. Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)

1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.



```{r set up library}

# load libraries
library(raster)
library(rhdf5)
library(rgdal)

```

# Import Raster

```{r import-raster}
# Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

chm<-raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

# quick sanity check
plot(chm)
```

# Remove zero's and Classify 

```{r set-zeroes-reclassify}
# histo with raw data
old.chm<-hist(chm)

#sanity check
plot(old.chm, main="Non-tidy CHM Data")

#check density plot
density(chm)


# re-classify zeroes to NA
chm[chm==0] <-NA
plot(chm)
hist(chm)

## classify to distinguish  low and tall trees
# create matrix for reclassification

#create string of values
# by creating a matrix
height.class.m<- c(0, 6, 1,
            6, 30, 2,
            30, 50, 3,
            50, 100, 4)

# reshape vector into a matrix

rclass.m<- matrix(height.class.m, 
         ncol=3, 
         byrow=TRUE)

rclass.m

# reclassify raster
reclassified.chm<-reclassify(chm, rclass.m)

plot(reclassified.chm, main="Reclassified CHM Raster")


# Plot Re-classified Raster
pdf(file="TEAK", width=6, height=7)


density(reclassified.chm, 
        main="Canopy heights at Teakettle site\n with bin cutoffs in red",
        xlab= "Height (m)")

abline(v= rclass.m[ , 2], col ="red") ## adds abline based on classification cutoffs

dev.off()
```

```{r}




# 1. Export the plot figure to a pdf – publishable

```{r}
pdf(file="TEAK_CHM_density_with_braeks.pdf", width=6, height=7)

```




```{r export-TIFF}
#1. Export the plot figure as a TIFF – publishable

writeRaster(reclassified.chm,
            filename="reclassified_CHM_TEAK_GEOTIFF.tif",
            format="GTiff",
            options="COMPRESS=LZW",
            overwrite = TRUE,
            NAflag = -9999)

```

