---
title: "Functionalizing Rasters from NEON"
author: "R. Remigio"
date: "June 21, 2016"
output: pdf_document
---

## Objectives

Start by putting up objectives/tasks that students will be working though:
1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

1. For the CHM, set values == 0 to NA (not trees)

1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that 6m is generally grasses / understory, 10m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize density and plot vertical cutoff lines.

1. Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)

1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.



```{r set up library}

# load libraries
library(raster)
library(rhdf5)
library(rgdal)

```

# Import Raster

```{r import-raster}
# Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

chm<-raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

plot(chm)
```

# Remove zero's and Classify 

```{r set-zeroes-to-NA-classify}

# histo with raw data
old.chm<-hist(chm)
plot(old.chm, main="Non-tidy CHM Data")

# re-classify zeroes to NA
chm[chm==0] <-NA

hist(chm)

## classify to distinguish  low and tall trees

#create string of values
# by creating a matrix
class.m<- c(0, 2, 1,
            2, 6, 2,
            6, 60, 3)

rcl.m<- matrix(class.m, 
         ncol=3, 
         byrow=TRUE)

rcl.m

new.chm<-reclassify(chm, rcl.m)

plot(new.chm, main="Reclassified CHM Raster")



```

# Plot Re-classified Raster

```{r plot-raster}
# make room for a legend
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
# plot
plot(new.chm,
     col=c("yellow","orange","tan"), main="Reclassifed CHM Raster" # hard code colors, unclassified (0)=white,
  	 #N (1) =blue, S(2)=green
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("low", "medium", "high"),  # make sure the order matches the colors, next
       fill = c("yellow", "orange", "tan"),
       bty="n") # turn off border

```

# 1. Export the plot figure to a pdf – publishable

```{r}

```



# 1. Export the plot figure to a TIFF – publishable




```{r export-TIFF}
#1. Export the plot figure as a TIFF – publishable

writeRaster(new.chm,
            filename="outputs/TEAK/Teak_nsAspect.tif",
            format="GTiff",
            options="COMPRESS=LZW",
            overwrite = TRUE,
            NAflag = -9999)

```

